### 2020-03-17 assignment_4/article.tex
✨ Add Problem 3 sketch response (w/o comments).

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\usepackage{etoolbox}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

\makeatletter
\renewcommand{\@seccntformat}[1]{}
\makeatother

\begin{document}

\title{\textbf{EECS 340: Assignment 4}}

\author{Shaochen (Henry) ZHONG, \ilc{sxz517} \\ Yuhui ZHANG, \ilc{yxz2052}}
\date{Due and submitted on 03/19/2020 \\ EECS 340, Dr. Koyut{\"u}rk}
\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 1}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{{fig/fig_1.png}}
    \caption{Decision Tree of \textsc{Quicksort}}
    \label{fig:1}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 2}

\subsection{Procedure \textsc{PreProcess}(A)}

\begin{algorithm}[H]
\caption{PreProcess(A)}\label{PreProcess}
    \begin{algorithmic}[1]
        \Procedure{PreProcess}{A, p, r}
        \State \text{Let $C[0, 1, ..., k]$ be an arry of $0$.}
        \For {$i \gets 0$ \textbf{to} $n$}
            \State $C[A[i]] = C[A[i]] + 1$
        \EndFor
        \For {$j \gets 1$ \textbf{to} $k$}
            \State $C[j] = C[j] + C[j-1]$
        \EndFor
        \State \Return $C$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsection{Procedure \textsc{Query}(A, a, b)}

\begin{algorithm}[H]
\caption{Query(A, a, b)}\label{Query}
    \begin{algorithmic}[1]
        \Procedure{Query}{A, a, b}
        \If {$a == 0$}
            \State \Return $A[b]$
        \Else
            \State \Return $A[b] - A[a-1]$
        \EndIf
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 3}

\begin{algorithm}[H]
\caption{Sparse-Transpose(R, C, V, m, n, k)}\label{transpose}
    \begin{algorithmic}[1]
        \Procedure{Sparse-Transpose}{A, a, b}
        \State \text{Let $CH[0, 1, ..., n]$ be an arry of $0$.}
        \State \text{Let $VH[0, 1, ..., n]$ be an arry of $0$.}
        \For {$i \gets 0$ \textbf{to} $n$}
            \State $start \gets R[i]$
            \State $end \gets R[i+1]-1$
            \State \text{Let $CPR[\ ]$ be an empty array.}
            \State \text{Let $VPR[\ ]$ be an empty array.}
            \For {$j \gets start$ \textbf{to} $end$}
                \State $CPR[i].append(C[j])$
                \State $VPR[i].append(V[j])$
            \EndFor
            \State $CH[i].append(CPR)$
            \State $VH[i].append(VPR)$
        \EndFor
        \State \text{Let $R[\ ]$ be an empty array.}
        \State \text{Let $C[\ ]$ be an empty array.}
        \State \text{Let $V[\ ]$ be an empty array.}
        \State $V.append(0)$
        \State \text{Let $len\_max$ to be the maximum length amoung all elements in $CH$}.
        \For {$i \gets 0$ \textbf{to} $len\_max$}
            \State $\textsc{Pop-Smallest}(R, C, V, CH, VH, len\_max)$
        \EndFor
        \For {$i \gets 1$ \textbf{to} $n$}
            \State $R[i] = R[i] + R[i-1]$
        \EndFor
        \State \Return R, C, V
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{Pop-Smallest(R, C, V, CH, VH, len\_max)}\label{pop}
    \begin{algorithmic}[1]
        \Procedure{Pop-Smallest}{R, C, V, CH, VH, len\_max}
        \State $min\_value \gets CH[0][0]$
        \State $min\_i \gets 0$
        \State $min\_j \gets 0$
        \For {$i \gets 0$ \textbf{to} $len\_max$}
            \State $R\_counter \gets 0$
            \For {$j \gets 0$ \textbf{to} $n$}
                \If {$CH[j][i] < min\_value$ or ($CH[j][i] == min\_value$ and $j < min\_j$)}
                    \State $min\_value \gets CH[j][i]$
                    \State $min\_i \gets i$
                    \State $min\_j \gets j$
                    \State \text{continue}
                \EndIf
                \If {$CH[j][i] == min\_value$}
                    \State $R\_counter \gets R\_counter + 1$
                    \If {$j < min\_j$}
                        \State $min\_value \gets CH[j][i]$
                        \State $min\_i \gets i$
                        \State $min\_j \gets j$
                    \EndIf
                \EndIf
            \EndFor
            \State $CH[min\_j].remove(min\_i)$
            \State $C.append(min\_j)$
            \State $V.append(VH[min\_j][min\_i])$
            \State $VH[min\_j].remove(min\_i)$
            \State $R.append(R\_counter)$
        \EndFor
    \end{algorithmic}
\end{algorithm}

% \section{References}
%
% \nocite{*}
% \raggedright
% \bibliography{references.bib}
% \bibliographystyle{plain}

\end{document}