### 2020-04-07 assignment_5/article.tex
✨ Add Problem 1(b) responses.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\usepackage{etoolbox}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

\makeatletter
\renewcommand{\@seccntformat}[1]{}
\makeatother

\begin{document}

\title{\textbf{EECS 340: Assignment 5}}

\author{Shaochen (Henry) ZHONG, \ilc{sxz517}}
\date{Due and submitted on 04/06/2020 \\ EECS 340, Dr. Koyut{\"u}rk}
\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 1}

I did \textbf{(ii)} and \textbf{(vi)} on \textbf{Problem 1(a)}.
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{(a)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\paragraph{(i)}

The last operation base on operator $o_{\text{last}}$ for $o_{\text{last}} \in \{o_1, o_2, o_3, ..., o_{n-1}\}$.

\paragraph{(ii)} Assuming the last operation will be $T[i, j] \ o_{\text{last}} \ T[j+1, k] = T_{\text{final}}$, where $T_{\text{final}}$ provides the optimal output for $x_1 o_1 x_2 o_2 ... x_{n-1} o_{n-1} x_n$. There must be:

\begin{itemize}
    \item $T[i, j]$ provides the optimal output for $x_i o_i x_{i+1} o_{i+1} ... x_{j}$.
    \item $T[j+1, k]$ provides the optimal output for $x_{j+1} o_{j+1} x_{j+2} o_{j+2} ... x_{k}$.
\end{itemize}

\begin{proof}
    Without lost of generality, assume $T[i, j]$ is not optimal for $x_i o_i x_{i+1} o_{i+1} ... x_{j}$, there must be an $T'[i, j]$ where $T'[i, j] > T[i,j]$. Then we may have $T'[i, j] \ o_{\text{last}} \ T[j+1, k] > T_{\text{final}}$. \\
    However, it is known that $T[i, j]$ provides the optimal output for $x_i o_i x_{i+1} o_{i+1} ... x_{j}$, for $T'[i, j] > T[i,j]$ to exist,  $T[i, j]$ is no longer optimal. Thus we may say $T[i, j]$ is optimal by contradiction.
\end{proof}

\paragraph{(iii)}

\begin{itemize}
    \item Subproblem: $T[i, j]$ for the optimal output of $x_i o_i x_{i+1} o_{i+1} ... x_{j}$.
    \item Overall problem: $T[1, n]$ for the optimal output of $x_1 o_1 x_2 o_2 ... x_{n-1} o_{n-1} x_n$.
\end{itemize}

\paragraph{(iv)}

\begin{equation}
    T[i,k] =
    \begin{cases}
        x_i & \text{if \ } i = k \\
        max\{T[i,j] \ o_j \ T[j+1, k]\} &\text{for \ } i \leq j \leq k \\
    \end{cases}\nonumber
\end{equation}

\paragraph{(v)}

We have $T(n) = \Theta(n^3)$. As there are $\Theta(n^2)$ subproblems and for each subproblem of $T[i,k]$, there are $\Theta(k-i)$ choices (which is essential $\Theta(n)$). Also each choice takes $\Theta(1)$ too execute. \\
$\Longrightarrow T(n) =  \Theta(n^2) \cdot  \Theta(n) \cdot \Theta(1) =  \Theta(n^3)$

\paragraph{(vi)}

\begin{algorithm}[H]
\caption{Largest-Output(T, n)}
    \begin{algorithmic}[1]
        \Procedure{Largest-Output}{T, n}
        \For {$i \gets 1$ \textbf{to} $n$}
        \EndFor
    \end{algorithmic}
\end{algorithm}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{(b)}

\paragraph{(i)} Determine the last coin $c_{\text{last}}$ which will be added to the change set $\{c_1, c_2, c_3, ..., c_{\text{last}-1}\}$. Where the sum of elements of such set plus $c_{\text{last}}$ should be equal to $t$.

\paragraph{(iii)}

\begin{itemize}
    \item Subproblem: $f(v)$ provides the minimum number of coins to make change for $v$, given $n$ types of coin denominations are available. There must be  $v \in [0, t]$ and $v$ must be equal to the sum of some coins from the given $n$ types of coin denominations, to represent a value that is achievable by given coins.
    \item Overall problem: $f(t)$ provides the minimum number of coins to make change for $t$, given $n$ types of coin denominations are available.
\end{itemize}

\paragraph{(iv)}

Let $n_c$ be a type of coin from the provided $n$ types of coins, which means $n_c$ can potentially hold $n$ types of values. We will try all $n_c$(s) as we don't know which coin to use for the $c_1$.

\begin{equation}
    f(v) =
    \begin{cases}
        0 & \text{if \ } v = 0 \\
        min\{f(v - n_c) + 1\} & \text{if \ } v > 0 \\
    \end{cases}\nonumber
\end{equation}

\paragraph{(v)}

We have $T(n) = \Theta(nt)$. There are $\Theta(t)$ subproblems, as we can divide at most $t$ groups of coin(s) for a total value of $t$ (as for all coins being value $1$), and for each group there are at most $n$ possible choices to iterate (thus $\Theta(n)$). Each choice takes $\Theta(1)$ to execute. \\
$\Longrightarrow T(n) =  \Theta(t) \cdot \Theta(n) \cdot \Theta(1) =  \Theta(nt)$

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{(c)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 2}

% \section{References}
%
% \nocite{*}
% \raggedright
% \bibliography{references.bib}
% \bibliographystyle{plain}

\end{document}