### 2020-04-28 assignment_7/article.tex
🐛 Fix incorrect Problem 2(a) due to misread of uv being an edge.

\documentclass[11pt]{article}
\usepackage{setspace}
\setstretch{1}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb, amsthm}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[hang, flushmargin]{footmisc}
\usepackage[colorlinks=true]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{footnotebackref}
\usepackage{url}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{inconsolata}
\usepackage[papersize={8.5in,11in}, margin=1in]{geometry}
\usepackage{float}
\usepackage{caption}
\usepackage{esint}
\usepackage{url}
\usepackage{enumitem}
\usepackage{subfig}
\usepackage{wasysym}
\newcommand{\ilc}{\texttt}
\usepackage{etoolbox}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{matrix,positioning,arrows.meta,arrows}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

\makeatletter
\renewcommand{\@seccntformat}[1]{}
\makeatother

\begin{document}

\title{\textbf{EECS 340: Assignment 7}}

\author{Shaochen (Henry) ZHONG, \ilc{sxz517} \\ Yuhui ZHANG, \ilc{yxz2052}}
\date{Due and submitted on 04/27/2020 \\ EECS 340, Dr. Koyut{\"u}rk}
\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 1}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{(a)}

Set the each team as a node (vertex) in graph $G$ and make every node fully connected (representing the fact that all teams have played against each other). If one team $A$ beats another team $B$, we will have a directed edge of $A \rightarrow B$ betwen the nodes representing $A$ and $B$.

\begin{algorithm}[H]
\caption{DFS(G)}
    \begin{algorithmic}[1]
        \Procedure{DFS}{G}
        \For{each $u \in G.V$}:
            \State $u.color$ = WHITE
            \State $u.z$ = 0
        \EndFor
        \For{each $u \in G.V$}:
            \If{$u.color$ == WHITE}
                \State \textsc{DFS-Game-Visit(G, u)}
            \EndIf
        \EndFor
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{DFS-Game-Visit(G, u)}
    \begin{algorithmic}[1]
        \Procedure{DFS-Game-Visit}{G, u}
        \State $z = NULL$
        \State $u.color$ = GRAY
        \For{each $v \in Adj(u)$}:
            \If{$v.color$ == WHITE}
                \State $z = v.rank$
                \State $z = \min(z, \textsc{DFS-Game-Visit(G, v)})$
            \EndIf
        \EndFor
        \State $u.color$ = BLACK
        \If{$z == NULL$}
            \State $u.z = 0$
            \State $z = u.rank$
        \Else
            \State $u.z = z$
        \EndIf
        \State \Return $z$
    \end{algorithmic}
\end{algorithm}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{(b)}

\paragraph{Justification for runtime} The algorithm is $O(m+n)$ as DFS will first traverse every node, which means it will at least be $O(n)$. In addition, since the graph $G$ is implemented using adjacency list, for each node we will have to traverse through all its adjacent edges -- where such number can be at most $m$ (total number of edges in $G$) for a single node. Thus, the total time complexity is $O(m + n)$.

\paragraph{Justification for correctness} The algorithm basically performs a DFS travesal on the graph $G$ and set the node \ilc{u} with smallest \ilc{rank} value as \ilc{z} to all of node \ilc{u}'s ancestor nodes. Due to the nature of DFS, node \ilc{u}'s ancestor nodes can legally have \ilc{u.rank} as their domination factor, as being \ilc{u}'s ancestors imply the fact these teams have won against team \ilc{u}. Also in \texsc{DFS-Game-Visit} we will not set a node's \ilc{.z} value unless such node is marked as \ilc{BLACK} (which means all of its decedents have been explored). Combine these two observations, each node has explored all of its decendents nodes, and the node's decendent node with the smallest \ilc{rank} will be assigned as all its ancestor nodes' \ilc{.z} value. The algorithm is a perfect mimic of the game logic of domination factor and guaranteed to be correct.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 2}

Note the tuple below each node represents their \ilc{(start-time, finish-time)}.

\subsection{(a)}

Proof with cases:

There are two possible cases where we can have a $v.f > u.f$:

\begin{itemize}
    \item Case 1: $v.d < u.d < u.f < v.f$\newline
    In this case $u$ is a decent of $v$, which means there must be a path of $v \rightarrow u$; and known that we have $uv \in E$ (from the question instruction), this implies there is a direct connection of $u \rightarrow v$. Thus, now we have a path of $v \rightarrow u$ and a path of $u \rightarrow v$, $uv$ is guaranteed to be on a circle.
    \item Case 2: $u.d < u.f < v.d < v.f$\newline
    In this case we have $u$ completely discovered before $v$. This suggest there shoule be no relationship between $u$ and $v$, as $v$ should not be a decendent nor ancestor of $u$. However, it is known from the question that we have $uv \in E$, which implies $v$ must be a decendent of $u$. Thus, this case will not be valid under the question restriction and therefore disregarded.
\end{itemize}

As all legal case(s) shows $uv$ is guaranteed to be on a circle, we may prove the statement to be true.

\subsection{(b)}

Disprove with conterexample:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.35\linewidth]{{fig/fig_2_b.png}}
\end{figure}

It is shown that there is a path from $v$ to $u$ as $v \rightarrow a \rightarrow u$, however $uv$ is still a cross edge as $v.d < v.f < u.d < u.f$ ($2 < 3 < 4 < 5$). Thus, this conterexample disproves the statement.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Problem 4}

\subsection{Shaochen (Henry) ZHONG's proof of course evalution}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\linewidth]{{fig/fig_4_a.png}}
\end{figure}

\subsection{Yuhui ZHANG's proof of course evalution}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\linewidth]{{fig/fig_4_b.png}}
\end{figure}

% \section{References}
%
% \nocite{*}
% \raggedright
% \bibliography{references.bib}
% \bibliographystyle{plain}

\end{document}